Brownian dynamics simulation of cytoplasmic dynein's powerstroke reproduces bidirectional stepping

### 2019-01-16 papers/paper/paper.tex
Add very rough additions to introduction

Dynein's mechanism of motion generation is the least understood of all motor proteins. Kinesin and myosin both take consistent steps of the same size \cite{kinesin-step-size, myosin-step-size}. In contrast, dynein is known to take highly variant steps both forwards and backwards, but averaging 16nm \cite{yildizpaper, weihongpaper}. Another interesting feature of dynein is the 20nm separation between its ATP hydrolysis and the site and MT binding site \cite{3vkh-cite}. The protein must transmit information about ATP binding across this distance in order to coordinate its steps \cite{mt-atp-coupling}. Explaining these phenomena is necessary to fully understand dynein's mechanism.

Studies over the past 15 years have shed much light on dynein's stepping mechanism. The motor is known to switch between two conformational states depending on its ATP hydrolysis state \cite{burgess-paper, FRETstatepaper, carter-paper, nicastro, schmidt-carter}. It is also known that ATP binding can induce structural changes within the motor which increase unbinding rate \cite{leschziner, carter-paper}, and similarly that microtubule rebinding can influence nucleotide unbinding rate \cite{mt-atp-coupling}. Taken together this evidence is synthesized into a powerstroke model \cite{cianfroccoreview, imamulamodel, tsygankovscheme}. In the powerstroke model, dynein binds ATP, causing structural changes which eventually lead to microtubule unbinding at its MTBD. These structural changes transition the motor into a more open position, causing the MTBD to diffuse forwards on the microtubule. This biased search is known as the prestroke. Eventually the MTBD rebinds the microtubule, causing structural changes inducing the ejection of hydrolyzed ADP and transitioning back to the closed position. This causes the motor to lurch forward in a process called the powerstroke. The protein is now ready to re-bind ATP and take another step.

Ample evidence exists that dynein engages in an ATP-induced prestroke and MT-binding-induced powerstroke. However, to our knowledge, no work has been done to test whether a structural model of dynein obeying the proposed mechanochemical cycle is sufficient to reproduce experimental stepping data. Existing computational models use chemical rate transitions and assume independence of steps, without simulating the entire mechanochemical cycle \cite{sarlahmodel}. No models have combined structural and kinetic data to simulate a dynein model in a biophysically realistic way. It is necessary to do so to properly test the powerstroke hypothesis. Such a study would determine if the powerstroke mechanism is sufficient to explain dynein's walk, or if further explanation is necessary. Others have theorized that elements such as electrostatic interactions with the microtubule \cite{longrangemt}, tension-gating between the two heads \cite{yildizcleary, imaiburgess}, ring stacking \cite{sarlahmodel}, discrete microtubule binding sites \cite{trottmodel}, or elasticity of the stalk \cite{burgess-paper} are important for dynein's walk. Such a study would be able to find the minimum criteria for generating dynein's motion.

Here we present a course-grained model of cytoplasmic dynein which treats its domains as connected rigid bodies. Chemical transitions and conformational changes consistent with the powerstroke model are accounted for via elastic forces. The dynamics are determined using Brownian dynamics. The model is capable of generating directed motility at rates similar to experiment. Further, we show that the model displays more complex behavior, such as tension gating and force-mediated bidirectionality.

%% just say that we included tension gating

@@

\subsection*{Unbinding}
Unbinding occurs when a bothbound MTBD unbinds from the microtubule. This event can occur for either MTBD, and each domain has its own unbinding rate $k_{ub}$. To account for the tension-gating proposed by Yildiz \textit{et al} \cite{yildiz} to allow for interhead communication and bias the motor forward, the parameter $c$ is used to alter the unbinding rates of either MTBD. For domain $i \in \{n, f\}$ with binding domain angle $\theta_{ib}\left(\theta_{nm}, \theta_{fm}\right)$, unbinding rate is given by $k_{ub} = k^0_{ub}e^{-c\left(\theta_{ib}-\theta^{eq}_{ib}\right)}$. This will bias the model to unbind more preferentially for the lagging MTBD due to a higher likelihood for $\theta^{eq}_{ib} > \theta_{ib}$ for the lagging head. Values used for parameters $k^0_{ub}$ and $c$, along with $k_b$, are shown in Table \ref{table:params}.

>>>>

Dynein's mechanism of motion generation is the least understood of all motor proteins \cite{}. Kinesin and myosin both take consistent steps of the same size \cite{kinesin-step-size, myosin-step-size, is-this-true-in-general-i-thought-only-types-of-myosin}. In contrast, dynein is known to take highly variant steps both forwards and backwards, averaging 16nm \cite{yildizpaper, weihongpaper}. Another interesting feature of dynein is the 20nm separation between its ATP hydrolysis and the site and MT binding site \cite{3vkh-cite}. The protein must transmit information about ATP binding across this distance in order to coordinate its steps \cite{mt-atp-coupling}. Explaining these phenomena is necessary to fully understand dynein's mechanism.

Studies over the past 15 years have shed much light on dynein's stepping mechanism. The motor is known to switch between two conformational states depending on its ATP hydrolysis state \cite{burgess-paper, FRETstatepaper, carter-paper, nicastro, schmidt-carter}. It is also known that ATP binding can induce structural changes within the motor which increase unbinding rate \cite{leschziner, carter-paper}, and similarly that microtubule rebinding can influence nucleotide unbinding rate \cite{mt-atp-coupling}. Taken together this evidence is synthesized into a mechanochemical cycle \cite{cianfroccoreview, imamulamodel, tsygankovscheme}. In this scheme, dynein binds ATP, causing structural changes which eventually lead to microtubule unbinding at its MTBD. These structural changes transition the motor into a more open position, causing the MTBD to diffuse forwards on the microtubule. This biased search is known as the prestroke. Eventually the MTBD rebinds the microtubule, causing structural changes inducing the ejection of hydrolyzed ADP and transitioning back to the closed position. This puts the motor in the post-stroke state, moved forward slightly. The protein is now ready to re-bind ATP and take another step.

This mechanochemical cycle of chemical transitions is well supported by evidence \cite{citemore}. However, there is still relatively little known about the dynamic mechanism of domain motion dynein uses to achieve its forward motion by transitioning between these states \cite{lippert}. There exist models theorizing how dynein's domains move during its forward step, such as the winch \cite{} and powerstroke \cite{} models, which can reproduce experimental data \cite{}. However, to our knowledge, no studies exist which study if either of these dynamic mechanisms is feasible for the dynein motor to produce. Existing computational models of walking use chemical rate transitions and assume independence of steps, without simulating the precise protein dynamics \cite{}. Molecular dynamics simulations exist, but are of too short a timescale to describe dynein's forward stroke \cite{}. It is necessary to model dynein dynamically at a large timescale to fully understand how feasible models like the winch and powerstroke are. Further, such long-timescale dynamic models could be used to find what elements of dynein are sufficient to produce forward motion. Further, an accurate dynamical model would allow us to test if certain dynamical elements, such as electrostatic interactions with the microtubule \cite{longrangemt}, ring stacking \cite{sarlahmodel}, discrete microtubule binding sites \cite{trottmodel}, or elasticity of the stalk \cite{burgess-paper} are necessary for dynein's walk.

Here we present a course-grained model of cytoplasmic dynein which reduces the mechanochemical cycle down to pre- and post-stroke states. We use cryo-EM data to create rough structural models of these conformers, and impose inter-domain tension and elastic restoring forces to describe the dynamics within each conformer. We then model the state transition as a transition between two equilibria, and use Brownian dynamics to simulate the model's behavior in physically realistic drag and diffusion conditions. This minimal model spontaneously generates a powerstroke-like dynamic mechanism, and uses it to produce bidirectional motion. The model also reproduces several experimental measures, such as step size, velocity and force-mediated directionality control.

@@

\subsection*{Unbinding}
Unbinding occurs when a bothbound MTBD unbinds from the microtubule. This event can occur for either MTBD, and each domain has its own unbinding rate $k_{ub}$. To account for the tension-gating proposed by Yildiz \textit{et al} \cite{yildiz, yildizcleary} to allow for interhead communication and bias the motor forward, the parameter $c$ is used to alter the unbinding rates of either MTBD. For domain $i \in \{n, f\}$ with binding domain angle $\theta_{ib}\left(\theta_{nm}, \theta_{fm}\right)$, unbinding rate is given by $k_{ub} = k^0_{ub}e^{-c\left(\theta_{ib}-\theta^{eq}_{ib}\right)}$. This will bias the model to unbind more preferentially for the lagging MTBD due to a higher likelihood for $\theta^{eq}_{ib} > \theta_{ib}$ for the lagging head. Values used for parameters $k^0_{ub}$ and $c$, along with $k_b$, are shown in Table \ref{table:params}.
